MSD panels 1, 2 and 3 Explant Sup

### 2016-08-22 MSD_panels_1_2_3_explant_sup_analysis.Rmd
fixed greek letters, changed detection range plot

We used 3 different MSD panels containing a total of 20 analytes analyzed for 46 samples(see caveat). Bsaed on results from running test samples, we decided to dilute all the supernants 1:5 with MSD's diluent 43 before running the plates. All samples (except 1, see caveats) were run in duplicate.

@@

Due to a possible pipetting error, we were not able to use the data from the standards that we ran for panel 1. Instead, we used the standards from a different run of the same panel to calculate concentrations for the experimental samples(see explant_sup_README for details).

@@

@@

```{r panel 2 and 3 standards, echo=FALSE, message = FALSE, warning = FALSE}

@@

* Standards with values below the detection range (<2.5x the standard deviation of the lowest standard) are give unreliable concentrations, so I did not include those in the plots.

* When standards are below the curve fit (outside the range of the standard curve), the software does not calculate concentrations so you can't get a percent recovery.

@@

`r pander(explant_stds_all_panels_below)r`

@@

#What number of samples are in each category of detection?
explant_all_panels_noStds_Below <- explant_all_panels_noStds%>%
 group_by(Assay, Detection.Range)%>%
  summarise(Number_of_Samples = n())

@@

ggplot(explant_all_panels_noStds_Below, aes(x = Assay, y=Number_of_Samples))+
         geom_point(aes(color = Detection.Range),size = 2)+
  theme(axis.text.x = element_text(angle=25))+
  ggtitle("Number of Samples in each Detection Range per Analyte")

@@

Also note:
Some of the samples had one rep that was below detection and another that was above. For these samples I used the remaining rep as the "average" for that sample.
```{r correcting analyte names, echo=FALSE, message = FALSE, warning = FALSE}

#excel messes up the greek letters of the analytes so I'm fixing them here.

#unicode greek lower case alpha = U03B1

@@

#make a list of the analytes
analyteList<-as.list(levels(avgConc_explant_all_panels_noStds_noBelow$Assay))

@@

lapply(analyteList, FUN = plotAnalytes)
```

>>>>

We used 3 different MSD panels containing a total of 20 analytes analyzed for 46 samples (see caveat). Based on results from running test samples, we decided to dilute all the supernants 1:5 with MSD's diluent 43 before running the plates. All samples (except 1, see caveats) were run in duplicate.

@@

Due to a possible pipetting error, we were not able to use the data from the standards that we ran for panel 1. Instead, we used the standards from a different run of the same panel to calculate concentrations for the experimental samples (see explant_sup_README for details).

@@

#fixing the greek letters!!
explant_all_panels$Assay <-recode(explant_all_panels$Assay,
                                      "IFN-α2a" = paste("IFN-\U03B1","2a", sep = ""),
                                      "IL-1α" = "IL-1\U03B1",
                                      "MIP-1α"="MIP-1\U03B1",
                                      "TNF-α" = "TNF-\U03B1",
                                      "MIP-3α" = "MIP-3\U03B1",
                                      "MIP-1β" = "MIP-1\U03B2",
                                      "IL-1β" = "IL-1\U03B2",
                                      "IFN-γ" = "IFNg")

@@

```{r standards, echo=FALSE, message = FALSE, warning = FALSE}

@@

* Standards with values *below the detection range* (<2.5x the standard deviation of the lowest standard) give unreliable concentrations, so I did not include those in the plots.

* When standards are *below the curve fit* (outside the range of the standard curve), the software does not calculate concentrations so you can't get a percent recovery.

@@

`r pander(explant_stds_all_panels_below)`

@@

#What number of samples are in each category of detection?
explant_all_panels_noStds_Below <- explant_all_panels_noStds%>%
  mutate(Panel = ifelse(str_detect(Plate.Name, "panel_1"),"Panel 1",ifelse(str_detect(Plate.Name,"panel_2"),"Panel 2","Panel 3")))%>%
 group_by(Panel,Assay, Detection.Range)%>%
  summarise(Number_of_Samples = n())

@@

ggplot(explant_all_panels_noStds_Below, aes(x = Assay, y=Number_of_Samples))+
         geom_point(aes(color = Detection.Range),size = 2)+
  theme(axis.text.x = element_text(angle=25))+
  labs(y = "Number of Samples", x = "Analyte")+
  facet_wrap(~Panel, scales = "free")+
  ggtitle("Number of Samples in each Detection Range per Analyte")

@@

Also note:
Some of the samples had one rep that was below detection and another that was in range. For these samples I used the remaining "in range" rep as the "average" for that sample.

@@

#make a list of the analytes

#I'm using a unique vector here instead of the levels of Assay because I don't want to include IFNa2a OR Il-7. There weren't any data points in the detection range for those analytes (so I don't want to try to make a plot of it)

analyteList<-as.list(unique(avgConc_explant_all_panels_noStds_noBelow$Assay))

@@

lapply(analyteList, FUN = plotAnalytes)
